### 2019-03-04 report/final-report.tex
Checkpoint new section 3

@@

Our approach combines two strategies.  We use \textit{domain-specific programming
  language}, specifically, the Boomerang~\cite{boomerang, Matching10} language,
which is designed for writing bidirectional string transformations.  We also exploit
\textit{type- and example-directed program synthesis}, which uses enumerative
techniques to search a space of candidate solutions.  The search is constrained
by the type information about the program, as well as by example instances
provided by the user.

\subsection{Lenses and lens domain-specific programming languages}

@@

@@

During the course of this project, we have designed, analyzed and
implemented algorithms that demonstrate it is possible to synthesize
three classes of bidirectional transformations: (1) pure bijective
transformations, (2) bijections modulo equivalences classes (quotient
lenses) and (3)
bijections modulo projections (symmetric lenses).  The synthesis algorithms we have
designed take inputs that include a format specification (which
includes specification of equivalence classes) and a collection of
examples.  As the class of transformations becomes richer, the number
of potential programs grows dramatically.  As a result, the
corresponding inference algorithm slows and its ability to guess the
transformation desired by the user decreases, leading to reduced
accuracy.  We found that it was possible to overcome such challenges
through new heuristics that use information theory to help guide the
search for program transformations.  In addition, we found the use of
compositional synthesis, which is the process of breaking down a
complex synthesis task into smaller, more mangeable subtasks critical
when attempting to scale our prototype system up to be able to handle
larger transformation task.

>>>>

% Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.2,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 240}
\definecolor{vlightyellow}{RGB}{255, 255, 245}
\definecolor{lightgreen}{RGB}{235, 255, 240}
\definecolor{vlightgreen}{RGB}{247, 255, 250}
\definecolor{lightblue}{RGB}{245, 255, 255}
\definecolor{vlightblue}{RGB}{250, 255, 255}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

@@

\section{Lenses}\bcp{Is a separate section the best place for this?  Is
  this the kind of example we had in mind?  Is it OK that I cannibalized the
later subsection on lenses?  (We could alternatively move all this there.)}

\begin{figure}
\small
  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily,
xleftmargin=0pt,
]
Jane Doe: 38000
John Public: 37500
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingA}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
xleftmargin=0pt,
stringstyle=\upshape\sffamily
]
FirstLast,Company
Jane Doe,Healthcare Inc.
John Public,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[xleftmargin=0pt]
let salary = number | "unk"
let emp_salary =   name . " "
                 . name . ": " 
                 . salary
let emp_salaries = "" 
                 |   emp_salary
                   . ("\n" emp_salary)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingC}{\box0}

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[xleftmargin=0pt]
let company = (co_name . ("Co." | "Inc." | "Ltd.")) 
            | "UNK" 
let emp_ins = name . " " . name "," company 
let header = "FirstLast,Company" 
let emp_insurance = header . ("\n" . emp_ins)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightblue}}c|}
      \hline
      Management's data & HR's data \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}Management's type & 
      \cellcolor{lightblue} HR's type\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightblue}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively.
  }
  \label{fig:minimized-representations}
\end{figure}

Before surveying our methods and results on synthesis, we quickly sketch an
example of what a lens looks like, using a simplified situation drawn from a
hypothetical U.S. company. In this company, management and human resources
(HR) store information about employees in separate text files: management
stores the names of employees and their salaries while HR stores the names
of employees and their health insurance
providers. Figure~\ref{fig:minimized-representations} gives examples of the
two file formats and regular expressions describing them.

The company uses a simple symmetric lens to keep these files
synchronized. When management adds a new employee, say ``Chris Roe: 32500'',
this lens 
adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The default value
``UNK'' represents the fact that the employee's insurance company is currently
unknown. A similar update happens if HR adds a new employee before
management knows about them,
in which case the sentinel value ``unk'' reflects unknown salary information
on management's side.
Furthermore, if HR corrects an error in an employee's name, say changing ``John
Public'' to ``Jon Public'', the lens mirrors this change into management's
file. Not all the data is mirrored, however: management's file is not updated
in response to insurance changes, and HR's is oblivious to salary
changes. Simple symmetric lenses are appropriate for keeping these two files
synchronized because they contain a mix of shared and unshared information.

%Consider the previously
%described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
%(scheduled.job) synchronized. These are each formats specifying when to run
%recurring tasks. For the sake of exposition, we use simplified versions crontab
%and scheduled.job formats, with examples and corresponding regular expressions
%shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
%of tasks set to run at specific times every hour. In both formats, users
%can either provide specific minutes at which they want the program to run, or they
%can specify that the program should run at every minute.
%
%We wish for the jobs to be synchronized across our Windows and Linux devices.
%However, the commands that need to be run are typically different for each
%device. Paths are different in Windows and Linux, and, while there is typically a
%Windows equvialent for a Linux command, that equivalent command is not always
%named the same thing. So, while we want to be able to have the same number of
%commands, with the commands running at the same times, we don't want to
%align the actual commands that are run.

In principle, programmers can define lenses just by writing four arbitrary
functions by hand and showing they satisfy the round-tripping laws, but this
can be tedious and error prone.  A better idea is to provide a set of
combinators---a domain-specific language---for building complex, law-abiding
lenses from simpler ones.

@@

If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
$\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
stochastic regular expressions, so we mark plain REs with overbars.)
We illustrate some of these
combinators by defining lenses on subcomponents of the employee data formats.

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\BRegex$ and propagates data unchanged
in both directions.
%
\begin{lstlisting}
id(name)  :  name $\Leftrightarrow$ name
\end{lstlisting}
%
The identity lens moves data back and forth from source to target without
changing it. Both the \CreateR and \CreateL functions are the identity function
($\CreateR \App s = s$), and the put functions merely return the first argument
($\PutROf{\String}{\StringAlt} = \String$).

In contrast, $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ does not propagate any data at all
from one format to the other.  The \Disconnect lens takes four arguments: two
regular expressions ($\Regex, \RegexAlt$) and two strings ($\String, \StringAlt$). The regular expressions specify the formats
on the two sides, while the strings provide default values. 
%
\begin{lstlisting}
disconnect(salary, "", "unk", "")  :  salary $\Leftrightarrow$ ""
\end{lstlisting}
%
On creates, the input values are thrown away, and default values are
returned ($\CreateLOf{\StringAlt} =
\lstinline{"unk"}$), and on puts, the second argument is used and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). For 
example, the \lstinline{salary} field is only present in management files,
so the disconnect lens can ensure salary edits do not cause updates to the
HR file.
With the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
and \PutR will always return \lstinline{""}.

The insert lens \lstinline{ins} and the delete lens \lstinline{del} are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source or target format.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}
The \lstinline{ins} lens inserts a constant string when going from
left to right, while \lstinline{del} inserts a string when going from right
to left.

Finally, there are a number of combinators that construct
more complex lenses from simpler ones. These include concatenation
($\ConcatLensOf{\Lens_1}{\Lens_2}$, often written $\Lens_1 . \Lens_2$), variation
($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
For example, we could use the lens
%
\begin{lstlisting}
id(name) . id(" ") . id(name) . del(": ") . ins(",")
\end{lstlisting}
%
to transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
the left-to-right direction.

% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens is useful for synchronizing a series of items or rows in a
table. For example given a
lens \lstinline{employee_lens} that synchronizes data for a single employee, the
lens
%
\begin{lstlisting}
iterate(id("\n") . employee_lens)  :  ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%
transforms a list of employees in employees in the Management format to a list
of employees in the HR format and vice versa. These combinators
are combined in figure~\ref{fig:example_lens} to construct a complete lens
between the employee formats.

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens that synchronizes management and HR employee files}
  \label{fig:example_lens}
\end{figure}

Our approach to lens synthesis combines two strategies.  First, as
illustrated in the previous section, we use \textit{domain-specific
  programming language}, specifically the Boomerang  language~\cite{boomerang,
  Matching10}, which is designed for writing bidirectional string
transformations.  Second, we exploit \textit{type- and example-directed
  program synthesis}, which uses enumerative techniques to search a space of
candidate solutions.  The search is constrained by the type information
about the program, as well as by example instances provided by the user.

% \subsection{Lenses and lens domain-specific programming languages}

@@

In this project, we designed, analyzed and implemented algorithms that
synthesize three classes of bidirectional transformations: (1) pure
bijective transformations, (2) bijections modulo equivalences classes
(quotient lenses) and (3) bijections modulo projections (simple symmetric
lenses).  Our synthesis algorithms take as inputs a format specification
(which includes specification of equivalence classes) and a collection of
examples.  As the class of transformations becomes richer, the number of
potential programs grows dramatically.  As a result, the corresponding
inference algorithm slows and its ability to guess the transformation
desired by the user decreases, leading to reduced accuracy.  We overcame
these challenges through new heuristics that use information theory to help
guide the search for program transformations.  In addition, to scale our
prototype to larger transformation tasks, we found it necessary to work
compositionally, by breaking down complex synthesis tasks into smaller, more
mangeable subtasks.
